Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-12 main.tex
Updates from Overleaf

% which one is an empirical question - maybe hint that quadratic may result from RW on growth rate?

To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations. We consider a Poisson distribution where $\sigma^2 = \mu$, a geometric distribution where $\sigma^2 = \mu + \mu^2$, and a negative binomial distribution where the size parameter $\psi(\mu)$ depends on the mean $\mu$ such that $\sigma^2 = \mu + \mu^{2.5}/1000$. We see that when applying the CRPS on the natural scale, the expected CRPS grows in $\mu$ in all cases. When applying it after a log-transformation, the expected CRPS is almost independent of $\mu$ for the geometric distribution and large $\mu$, while smaller targets have higher expected CRPS in the Poisson case. When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution and grows for the two other distributions. 
%For the negative binomial distribution, targets with very small and very high expectation yield the highest expected CRPS. \textbf{Here it would be nice to also add a sentence on what happens under sqrt transformation.} % scores are independent of the mean of the forecast target. For a Poisson-distribution with 
% $\sigma = \mu$, scores on the log scale decrease as the mean increases. The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation .

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We generated 10,000 samples for three distributions with differing relationships between mean and variance. We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution. The first distribution (red) is a truncated normal distribution with constant variance (we chose a $\sigma = 1$ in order to only obtain positive samples). The second (green) is a geometric distribution (a negative binomial distribution with variance $\sigma^2 = \mu + \mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu + \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
\midrule
all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
\addlinespace
all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
\addlinespace
1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
\addlinespace
1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
\addlinespace
1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
\bottomrule
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}

@@

In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts. We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise. It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{Diks2011}). This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.

@@

@@

>>>>

To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations. We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with $\theta = 10$ and $\sigma^2 = \mu + 0.1\mu^2$, and a normal distribution with constant variance. We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster). The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two. When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance. When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two. As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution. 
    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots. Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}. Figure \ref{fig:XXX} shows the quality of the approximation in more detail. 
    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples). The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (left). B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19. See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}. Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
\midrule
all & all & -1.093 & 0.963 & 0.391 & 0.001 & -0.352 & 0.201\\
\addlinespace
all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.043 & 0.201\\
all & Deaths & -0.884 & 0.868 & 0.436 & -0.023 & 0.273 & 0.121\\
\addlinespace
1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.320 & 0.088\\
2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.112 & 0.164\\
3 & all & -1.001 & 0.984 & 0.410 & 0.008 & -0.094 & 0.241\\
4 & all & -0.757 & 0.986 & 0.469 & 0.015 & 0.000 & 0.299\\
\addlinespace
1 & Cases & -0.862 & 0.876 & 0.433 & -0.024 & 0.790 & 0.087\\
2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.109 & 0.238\\
4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.645 & 0.296\\
\addlinespace
1 & Deaths & -1.146 & 0.832 & 0.376 & -0.035 & 0.457 & 0.048\\
2 & Deaths & -0.981 & 0.867 & 0.416 & -0.028 & 0.443 & 0.084\\
3 & Deaths & -0.807 & 0.885 & 0.453 & -0.019 & 0.349 & 0.131\\
4 & Deaths & -0.602 & 0.891 & 0.501 & -0.011 & 0.125 & 0.194\\
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}

@@

In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts. We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise. It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{diksLikelihoodbasedScoringRules2011}). This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.

@@

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/SIM-score-approximation.png}
\caption{Comparison of expected CRPS values of an ideal forecaster and approximated values. Expected scores for three distributions as described in Section \ref{sec:methods:vst} and Figure \ref{fig:SIM-wis-state-size-mean} and approximated as described in \ref{sec:methods:vst}}
\label{fig:score-approx}
\end{figure}

@@

\begin{table}
\centering
\begin{tabular}{ccc}
\toprule
True value & \& & Median prediction\\
\midrule
$>0$ & \ & $>100\times$ true value\\
$>10$ & \  & $>20\times$ true value\\
$>50$ & \  & $<1/50\times$ true value\\
$= 0$ & \  & $>100$\\
\bottomrule
\end{tabular}
\caption{Criteria for removing forecasts. Any forecast that met one of the listed criteria (represented by a row in the table), was removed. Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters. When evaluating models against each other (rather than illustrating the effect of a transformation), one would prefer not to condition on the outcome when deciding whether a forecast should be taken into account. }
\label{tab:erroneous}
\end{table}
